\section{Introduction}

This thesis considers applying learning classifier systems (LCS's) to the prediction of time series data.
A time series as used here is a sequence of data successively measured through time.
Time series analysis encompasses many methods that attempt to understand such time series, aimed at either understanding the underlying theory present in the data points or to make real-world predictions.
Time series prediction is the use of a model to predict future events based on known past events: to predict future data points before they are measured.
One standard example is the opening price of a share of stock based on its past performance.

%\vspace*{-\baselineskip}    %You need to place this before a 
                            %subsection if it immediately follows
                            %section.
\subsection{Motivation}
No LCS to date has been designed for time series data but instead they were generally limited to Markov systems lacking any memory, which we viewed as a major limitation of LCS's.
LCS's are designed specifically with the concept of evolving an effective rule set for a specified problem, which is specifically the sort of capability that would be desirable for time series analysis and prediction: generating useful rule sets.

An LCS is an evolutionary algorithm that operates on a population comprised of rules referred to as the rule set: this rule set is used to attempt to classify a situation.
The first LCS was created by Holland \cite{HollandLCS} shortly after he created genetic algorithms (GA's) \cite{HollandGA}, one of the classical types of evolutionary algorithms.
Holland's first LCS originally used a GA as the evolutionary device.
Our system as described here also uses a GA for evolution, although it has been modified from the original form.

Holland's original LCS was quite complicated and failed to produce quality results for most real-world problems.
Because of this, the study of LCS's was somewhat inactive until Wilson introduced ZCS \cite{WilsonZCS}, a re-imagining of Holland's original LCS distilled to its most basic elements.
Wilson's ZCS was capable of producing acceptable results on certain problems and was simple enough to easily understand, reinvigorating LCS research.

A few years after introducing ZCS, Wilson modified it introducing XCS \cite{Wilson1995XCS}, which is currently one of the best performing and most popular LCS types.
Wilson's XCS was based on ZCS but with several important modifications mostly aimed at improving the accuracy of the rules produced and also for a more full coverage of the problem space by the rules.
A significant portion of the LCS's being worked on today are modifications or enhancements of Wilson's XCS.

One such enhancement of XCS is known as XCSR \cite{WilsonXCSR}, which was also developed by Wilson.
XCSR improves upon XCS by allowing it to operate with real-valued ranges for input instead of on the traditional ternary alphabet so common to LCS's, consisting of \emph{true}, \emph{false}, and a covering symbol (usually represented as $\#$ or $*$).

\input{background}
