\documentclass{article}

\usepackage{fullpage,longtable}

\title{Time Series Classifiers}
\author{Christopher Mark Gore\\
   chris-gore@earthlink.net\\
   (573)~452-3216}
\date{\today}

\begin{document}

\maketitle

\pagestyle{empty}
\thispagestyle{empty}

\section*{Abstract}
% An abstract answers the following questions:
% 1. What is the problem you are solving?
\noindent
A time series, as we use the term, is a sequence of data measured at successive time intervals.
Time series analysis refers to all of the methods employed to understand such data, either with the purpose of explaining the underlying system producing the data or to try to predict future data points in the time series.
Time series analysis is applicable to many problems, since there are so many areas that require a more thorough understanding of a time series or the prediction of future values of the time series.
The most typical historical examples of time series would be the weather and the financial markets, but there are many more real-world time series problems.

% 2. What is your approach?
We introduce here a new type of learning classifier system, originally derived from Wilson's XCSR, which allows for the modeling and prediction of one-dimensional time series data, and partially developed for use with multivariate times series.
This method works by modifying the makeup of the rules in the learning classifier system so that they are suitable for use on a time series, and therefore all of the operations (mutation, crossover, etc.) applied to the rules also were changed from their traditional forms.

\section*{The Stock Market}
\newcommand{\DJI}{\textasciicircum DJI}
\noindent
We experimentally determined many of the parameters that are best for use on the stock market.
We used actual historical data of the Dow Jones Industrial Average (\DJI), with daily trading data starting on August 20, 1990, with data ending on August 18, 2006.
The data was gathered from Yahoo! Finance.
The first 100 data points of the time series were skipped, allowing for historical data that far back even in the very first day of simulated data, causing an actual start of analysis of January 11, 1991.
In each of these experiments, a statistical sample of at least 30 runs was gathered, each run going on for 1,500 simulated trading days (2,167 actual days, 5.93 years), for an end of December 16, 1996.
At each trading day the stock was given the option to either put all of its resources into the \DJI\ or into a bank account yielding roughly 4\% per annum.
The system initially had \$1,000,000.00.

\section*{Current Results}
\newenvironment{cgoreErt}
{\begin{center}
   \begin{longtable}{c|lrrlr}
      B\&H & 806 & 53.733\% & \$2,745,309.50 & 1.0 & 18.54\%pa\\
      \hline
   \endfirsthead
      B\&H & 806 & 53.733\% & \$2,745,309.50 & 1.0 & 18.54\%pa\\
      \hline
   \endhead}
{\end{longtable}\end{center}}
\begin{cgoreErt}
arith mean & 765 & 50.985\% & \$2,090,409.40 & 0.76145 & 13.23\%pa \\
std dev & 22.8 & 1.521\% & \$295,592.78 & 0.10768 & 1.87\%pa \\
max & 817 & 54.467\% & \$2,848,646.30 & 1.03764 & 19.29\%pa \\
min & 725 & 48.333\% & \$1,536,175.30 & 0.55956 & 7.50\%pa
\end{cgoreErt}

\end{document}
